---
output: pdf_document
---

John Hendley and Sonish Lamsal

STA6375: Computational Statistics I

\begin{center}
Homework 4
\end{center}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse"); library("extrafont")
#theme_set(theme_minimal(base_family = "Roboto Condensed"))
theme_update(panel.grid.minor = element_blank())
```

## Question 1
### Part A
```{r num1a}
library("tidyverse"); library("extrafont")
theme_set(theme_minimal(base_family = "Roboto Condensed"))
theme_update(panel.grid.minor = element_blank())
ggplot(diamonds, aes(carat)) +
  geom_histogram(bins = 30) +
  geom_freqpoly(color = 'red', bins = 30) +
  scale_x_continuous(minor_breaks = NULL) +
  scale_y_continuous(minor_breaks = NULL) +
  theme_minimal()
diamonds
range(diamonds$carat)
```

### Part B
``` {r num1b}
x <- seq(0, 2*pi, pi/50)
y <- seq(0, 2*pi, pi/50)
df = data.frame(x, y)

d <- c()
for (x in 0:100) {
  for (y in 0:100) {
    d <- c(d, .5 * (sin(2*x*pi/50) + sin(2*y*pi/50)))
  }
}

df2 = data.frame(df, d)
#x <- "x" = 1:10, "y" = 1:10
#y <- seq(0, 2 * pi, pi / 6)
x <- c(0:2*pi)
y <- c(0:2*pi)
#fxy <- function(x,y) .5 * (sin(2*x) + sin(2*y))
#df <- expand.grid("x" = 0 : 6.5, "y" = 0 : 6.5)
#df <- tibble(x, y, d)
#fxy <- tibble(x, y, z = .5 * (sin(2*x) + sin(2*y)))
#df2 <- data.frame(fxy)
ggplot(df2, aes(x, y, fill = d)) +
  geom_density2d_filled() +
  geom_point() +
  theme_minimal() +
  scale_x_continuous(breaks = c(0, pi/2, pi, 3*pi/2, 2*pi),
    label = c(0, expr(pi / 2), expression(pi), expr(3 * pi / 2), expr(2 * pi)),
    minor_breaks = NULL) +
  scale_y_continuous(breaks = c(0, pi/2, pi, 3*pi/2, 2*pi),
    label = c(0, expr(pi / 2), expression(pi), expr(3 * pi / 2), expr(2 * pi)),
    minor_breaks = NULL) +
  coord_equal()
```


### Part C
``` {r num1c}
dens_df <- data.frame(x = 1:5, y = 1:5)
ggplot(dens_df) +
  geom_point(aes(x,y), color = "red", size = 4) +
  theme_minimal() +
  scale_x_continuous(limits = c(0, 6), minor_breaks = NULL) +
  scale_y_continuous(limits = c(0, 6), minor_breaks = NULL) +
  coord_equal()
```


### Part D
cut_interval() takes a lower and upper bound as well as number of groupings and divides the region up into the desired number of groupings. For example, cut_interval(10:20, 5) divides the interval into the following groupings: [10,12], (12, 14], (14, 16], (16, 18], (18, 20].

cut_number() takes two major arguments, x and n. The x is a vector that describes the upper and lower bound. The n is the number of groupings or bins. Also, the upper bound of x is the frequency that is divided into each of the bins. For example, cut_number(6:18, 3) creates 3 bins [6, 10], (10, 14], (14, 18] with 13 total frequency/objects in those bins (since there are 13 integers from 6 to 18). Those 13 objects are divided over the 3 bins so that the first gets 5 and the last two get 4.

cut_width also takes two key arguments, x and n. The x is again a vector that describes the upper and lower bound. But in this case the n is the width of each bin. Also, the upper bound of x is the frequency that is divided into each of the bins. For example, cut_width(3:11, 2) creates intervals of size 2 from 3 to 11: [3, 5], (5, 7], (7, 9], (9, 11]. with 9 total frequency/objects in those bins (since there are 9 integers from 3 to 11). Those 9 objects are divided over the 4 bins so that the first gets 3 and the last three get 2.


### Part E
``` {r num1e}
set.seed(2)
logistic <- function(x) 1 / (1 + exp(-x))
n <- 50
df <- tibble(
  x = c(rnorm(n, -1), rnorm(n, 1)),
  p = logistic(0 + 2*x),
  y = rbinom(2*n, 1, p)
)
ggplot(df, aes(x, jitter(y, amount = 0.03))) +
  theme_minimal() +
  geom_point() +
  geom_ribbon(aes(ymin = p - 0.1, ymax = p + 0.1),
    fill = "grey70", alpha = 0.6) +
  geom_line(aes(x, p), color = 'royalblue2', size = 1) +
  scale_x_continuous("x", limits = c(-4, 4), minor_breaks = NULL) +
  scale_y_continuous("y", limits = c(-.1, 1.1),
        breaks = c(0.00, 0.25, 0.50, 0.75, 1.00), minor_breaks = NULL)
```




#### Number 2
### Part A
```{r num2a}
dot <- function(x,y) {
  z <- c()
  for (i in 1:length(x)) {
    z <- c(z, x[i]*y[i])
  }
  z  
}

norm <- function(c) {
  total <- 0
  for (j in 1:length(c)) {
    total <- total + dot(c[j],c[j])
  }
  sqrt(total)
}

normalize <- function(v) {
  v / norm(v)
}

is_zero <- function(x) {
  norm(x) <= .Machine$double.eps
}

# Projects y onto x
project_onto <- function(y,x) {
  if (is_zero(x)) {
    c(0,0)
  } else {
  (dot(y,x)/norm(x)^2)*x
  }
}

#Made up an example for matrix v and A
(v <-  matrix(5:8, nrow = 2))
(A <- matrix (-5:-2, nrow = 2))
dif <- 1

# The while loop runs until the difference betwen consecutive iterations
# falls below a particular threshhold
while (dif > sqrt(.Machine$double.eps) ) {
  old_v <- v
  v <- normalize(v %*% A)
  v <- sign(v[1]) * v
  dif <- max(abs(v - old_v))
  print(v)
}

#Zeros out values that are effectively zero
(v <- replace(v, v < 10 ^ - 6, 0))


```



### Part B




### Part C
```{r num2c}
set.seed(2)
(A <- matrix(rpois(64, lambda = 5), nrow = 8))
```
